---
title: "Group 2W Analysis Report"
output: html_document
date: "2025-02-07"
editor_options: 
  markdown: 
    wrap: 72
---

# 1. R Setup

This section loads the required R packages and ensures all necessary
libraries are installed.

## 1.1 Load Required Packages

The following script checks if each package is installed; if not, it
installs and loads it.

```{r, include=TRUE, results='hide'}
packages <- c("dplyr", "ggplot2", "haven", "tidyr", "srvyr", "ipumsr", "survey", 
              "gtsummary", "car", "patchwork", "forcats", "broom", "purrr", 
              "see", "here", "bayestestR", "qqplotr", "plotly")

# Check and install missing packages
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

## 1.2 Data Setup

library(here) data_dir \<- here("data", "raw") \# 自动识别项目根目录下的
data/raw ddi \<- read_ipums_ddi(here(data_dir, "nhis_00007.xml")) data
\<- read_ipums_micro(ddi, data_file = here(data_dir, "nhis_00007.dat"))

```{r}
ddi <- read_ipums_ddi("/Users/recccs/Documents/GitHub/healthanalytics-project-group-W/data/raw/nhis_00007.xml"
)
data <- read_ipums_micro(ddi,data_file = "/Users/recccs/Documents/GitHub/healthanalytics-project-group-W/data/raw/nhis_00007.dat")

```

# 2.Data Cleaning

This section processes and cleans the dataset by renaming variables,
filtering out missing values, and recoding categories.

```{r}
data_clean <- data %>%
  mutate(
    AGE = YEAR - BIRTHYR,  # Calculate age
    WEIGHT = PERWEIGHT,     # Assign survey weights
  ) %>%
  filter(
    YEAR == 2015,             # Keep the only year 2015
    BIRTHAG > 10 & AGE > 10,  # Filter out impossible ages
    !MARSTCUR %in% c(0, 9),   # Remove unknown marital status
    !BIRTHAG %in% c(97, 98, 99),  # Remove unknown birth ages
    EDUCREC2 != 00,         # Remove unknown education levels
    RACENEW <= 900  # Remove unknown race cataories
  ) %>%
  mutate(  
    # Recode education levels
    EDUCREC2_group = case_when(
      EDUCREC2 == 10 ~ "None",  # Base category: Never attended/kindergarten only
      EDUCREC2 %in% c(20, 30, 31, 32) ~ "Primary",
      EDUCREC2 %in% c(40, 41) ~ "Secondary",
      EDUCREC2 %in% c(50, 51, 52, 53, 54, 60) ~ "Tertiary",
      TRUE ~ NA_character_  # Assign NA to all other categories
    ),
    # Create a Tertiary_dummy for later analysis
   Tertiary_dummy = 
    if_else(EDUCREC2_group == "Tertiary", 1, 0),
   
    # Recode region and race 
    REGION_group = case_when(
      REGION == 1 ~ "Northeast",            # Northeast (1)
      REGION == 2 ~ "North Central/Midwest", # North Central/Midwest (2)
      REGION == 3 ~ "South",                # South (3)
      REGION == 4 ~ "West",                 # West (4)
      TRUE ~ NA_character_                 # Any other value -> Other
    ),
    RACE_group = case_when(
      RACENEW == 100 ~ "White",
      RACENEW == 200 ~ "Black",
      RACENEW %in% c(300, 542) ~ "Native",
      RACENEW == 400 ~ "Asian",
      TRUE ~ "Others" # Assign all other other value to "Others"
    ),
    
    MARITAL_group = if_else(MARSTCUR %in% c(1, 2, 3), "Married", "Single")
   ) %>%
  filter(!is.na(EDUCREC2_group)) %>%  # Remove rows with NA in EDUCREC2_group
  mutate(
    # Set base groups for EDUCREC2_group, RACE_group, and REGION_group
    EDUCREC2_group = fct_relevel(EDUCREC2_group, "None"),  # Set "None" as the base for education
    RACE_group = fct_relevel(RACE_group,"White"),  # Set "White" as base for race
    REGION_group = fct_relevel(REGION_group,"Northeast"), # Set "Northeast" as the base for region
    MARITAL_group = fct_relevel(MARITAL_group,"Single")
  )

```

# 3.Survy Weight

This section sets up the survey design using sampling weights

```{r}
survey_design <- svydesign(
  ids = ~1, 
  weights = ~WEIGHT, 
  data = data_clean
)
```

# 4.Descriptive Statistics

Summarizes key demographic variables by education level.

```{r}
tbl_summary(
  data_clean,
  by = EDUCREC2_group,
  include = c(BIRTHAG, RACE_group, REGION_group,MARITAL_group),
  label = list(
    BIRTHAG ~ "Age at First Birth",
    RACE_group ~ "Race/Ethnicity",
    REGION_group ~ "Region",
    MARITAL_group~ "Maritial status"
  )
)

```

# 5. Regression Analysis

## 5.1 Baseline OLS Model (No Controls)

Examines the relationship between education and age at first birth
without additional covariates.

```{r}
model <-svyglm ( 
  BIRTHAG ~ EDUCREC2_group,
  design = survey_design
)

summary (model)

tbl_regression(model, exponentiate = FALSE, intercept = TRUE) %>%
  modify_table_body(~ .x %>% filter(label != "None")) %>%  # Remove "None" from EDUCREC2_group
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  bold_labels()

```

```{r}
p1 <- ggplot(data_clean, aes(x = BIRTHAG, fill = EDUCREC2_group)) +
  geom_density(alpha = 0.6) +
  labs(title = "Distribution of Age at First Birth",
       x = "Age at First Birth",
       y = "Density") +
  theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        text = element_text(family = "Times New Roman"),
        axis.title = element_text(size = 12),  
        axis.text = element_text(size = 12),
        legend.title = element_blank()
        )

p2 <- ggplot(data_clean, aes(x = EDUCREC2_group, y = BIRTHAG)) +
  geom_boxplot() +
  labs(title = "Effect of Education on Age at First Birth",
        x = "Education Level",
        y = "First Birth Age") +
  theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        text = element_text(family = "Times New Roman"),
        axis.title = element_text(size = 12),  
        axis.text = element_text(size = 12)
        )

p1/p2
```
## 5.2 Stepwise OLS Regression (Adding Controls)

Each model introduces additional controls. **Model 2: Adding Region
Control**

```{r}
model_2 <- svyglm(
  BIRTHAG ~ EDUCREC2_group + REGION_group,
  design = survey_design
)
summary(model_2)

tbl_regression(model_2, exponentiate = FALSE,intercept = TRUE) %>%
  modify_table_body(~ .x %>% filter(!label %in% c( "None","White","Northeast"))) %>% 
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  bold_labels()
```

**Model 3: Adding Race Control**

```{r}
model_3 <- svyglm(
  BIRTHAG ~ EDUCREC2_group + REGION_group + RACE_group,
  design = survey_design
)
summary(model_3)

tbl_regression(model_3, exponentiate = FALSE,intercept = TRUE) %>%
  modify_table_body(~ .x %>% filter(!label %in% c("Northeast","None","White"))) %>% 
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  bold_labels()
```
**Model 4: Adding Marital Status Control**

```{r}
model_4 <- svyglm(
  BIRTHAG ~ EDUCREC2_group + REGION_group + RACE_group + MARITAL_group,
  design = survey_design
)
summary(model_4)

tbl_regression(model_4, exponentiate = FALSE,intercept = TRUE) %>%
  modify_table_body(~ .x %>% filter(!label %in% c( "None","White","Single","Northeast"))) %>% 
  add_significance_stars(hide_ci = FALSE, hide_p = FALSE) %>%
  bold_labels()
```
